### 2019-03-19 writeup/paper.tex
added short corridor gridworld environment and updated paper

\documentclass{article}

\usepackage[english]{babel}
\usepackage[square, numbers]{natbib}
\usepackage{url}
\bibliographystyle{abbrvnat}
\setcitestyle{authoryear,open={(},close={)}}

\usepackage[autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{gensymb}
\usepackage{verbatim}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tikz-cd}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{centernot}
\usepackage{mathtools}
\usepackage[margin=1.0in]{geometry}
\usepackage{algorithm, algpseudocode}
\usepackage{bm}
\usepackage{caption}

\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepackage{tikz}
\usetikzlibrary{shapes}

\usepackage[title]{appendix}
\usepackage{cleveref}

\pgfplotsset{height=8cm, width=15cm,compat=1.9}

\DeclareMathOperator*{\esssup}{ess\,sup}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\Ex}{\mathbb{E}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

\newtheorem{definition}{Definition}
\newtheorem{exercise}{Exercise}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\title{Entropy and Information Content as Exploration Bonuses in REINFORCE}
\author{Alan Chan}
\date{}

\begin{document}
  \maketitle

  \section{Introduction}
  Policy gradient methods \citep[Chapter~13]{sutton2018reinforcement} in Reinforcement Learning (RL) are a way of learning a parametrized policy $\pi_\theta(a \mid s)$ in a Markov Decision Process (MDP) so as to maximize an objective function, usually the expected return $v_{\pi_\theta}(s_0)$ starting from a state $s_0$, where $v_{\pi_\theta}$ is the value function of $\pi_\theta$. In the following, we will assume that our state space $S$ and action space $A$ are both finite.

  A key component of policy gradient methods, and RL techniques in general, is sufficient exploration of the state-action space to identify optimal action(s) for a given state. One possible technique to improve such exploration is the use of entropy or information content as a pseudoreward signal. More formally, let $R(s, a)$ be a random variable denoting the reward obtained from following action $a$ in state $s$. We define the following two quantities:

  \begin{align*}
    R_e(s, a) &:= R(s, a) - \sum_{a' \in A} \pi_\theta(a' \mid s) \ln\left(\pi_\theta(a' \mid s)\right),
    R_i(s, a) &:= R(s, a) - \ln\left(\pi_\theta(a \mid s)\right),
  \end{align*}

  where $R_e(s, a)$ will be denoted the \textit{entropy pseudoreward}, and $R_i(s, a)$ will be denoted the \textit{information content pseudoreward}. Intuitively, for a given state-action pair $(s, a)$, $R_e(s, a)$ is maximized as a function of the policy $\pi_\theta$ when $\pi_\theta$ is a uniform distribution over all actions $a$. Thus, the reward is highest for a policy that uniformly selects actions at random, corresponding to a high degree of exploration. For a given state $s$, the maximum of $R_i(s, a)$ as a function of $a$ corresponds to a tradeoff between the maximum of $R(s, a)$ as a function of $a$ and $- \ln\left(\pi_\theta(a \mid s)\right)$ as a function of $a$. Intuitively, $- \ln\left(\pi_\theta(a \mid s)\right)$, the information content of taking action $a$, is at a maximum when $\pi_\theta(a \mid s)$ is at a minimum. In other words, the information content term in $R_i(s, a)$ encourages the agent to take actions that have low probability under the current policy, again corresponding to an increase in exploration.

  In this report, we focus upon the use of the entropy and information content pseudorowards in the REINFORCE algorithm \citep{williams1992simple}, pseudocode of which is in \citet[p.~328]{sutton2018reinforcement}. We evaluate compare both of the pseudorewards with the vanilla reward $R(s, a)$ to determine if any of the pseudorewards facilitate better exploration of the state-action space and thus whether a higher return is achieved in less time.

  \section{Experimental Details}
  \subsection{Algorithms}
  All code is available at \url{}.

  REINFORCE is the main algorithm used for this comparison. We chose REINFORCE because of its relative simplicity in the class of policy gradient algorithms. We chose not to use a baseline, as is common in practice to reduce the variance of REINFORCE updates, so that we may isolate the effects of the different reward functions.

  All methods were trained with SGD with step-sizes in the set $\{1 \cdot 10^{-4} \cdot 2^i : i \in \{0, 1, 2, 3, 4\}\}$. All algorithms were run for a set number of trials depending upon the environment, and the random seed was reset to the same value at the beginning of each set of trials for a given algorithm and environment to ensure fair comparison amongst all the algorithms.

  We plot learning curves of all the methods for each of our environments.

  \subsection{Environments}
  We performed experiments on three environments. The first environment is the short corridor gridworld from \citet[p.~323]{sutton2018reinforcement}.

  The second environment is cart pole,

  The third environment is mountain car

  \section{Results \& Discussion}

  \section{Conclusion}

  \bibliography{bibliography}
  \bibliographystyle{plain}
\end{document}